---
title: "TestRun"
output: pdf_document
---

This is my first try/test run to really bang through the code to make sure it works with a sample of Bradshaw district before I commit to using it for ALL the data across the three forests.

###0. Set-Up

Setting up the libraries I think I'll need for now.

```{r}
library(tidyverse)
library(readxl)
library(readr)
```

###I. Read In Data

Gotta bring in the data I need. I brought in the Frequency and Point Ground Cover data, not necessarily because I think I'll need it, but I might. Either way, good practice for me. VGS gives us a straight XLSX doc, which can be a bit unwieldy, so I'm going to pull specific sheets of data out, which should be specific to my protocol, and then convert them to CSVs.

```{r}
XLSXPath <- "RawDataTest/TestRunBradshaw.xlsx"

HerbOnlyDWR <- read_excel(XLSXPath, sheet = 1) #Herbaceous-Only Dry-Weight Rank
RegDWR <- read_excel(XLSXPath, sheet = 2) #Regular DWR
Freq <- read_excel(XLSXPath, sheet = 3) #Frequency
PGC <- read_excel(XLSXPath, sheet = 4) #Point Ground Cover
Cover <- read_excel(XLSXPath, sheet = "cover frequency (by quadrat) (c") #Canopy Cover-Frequency to be exact. Split between the PNF Gap protocol and the Tenth Acre protocol, so this will have to be broken down further.

#Reading all of the data into usable CSVs
write_csv(HerbOnlyDWR, "RawDataTest/HerbOnlyDWR.csv")
write_csv(RegDWR, "RawDataTest/RegDWR.csv")
write_csv(Freq, "RawDataTest/Freq.csv")
write_csv(PGC, "RawDataTest/PGC.csv")
write_csv(Cover, "RawDataTest/Cover.csv")
```

Looks like I'm getting warnings because of what looks like someone using a qualifier. Doesn't seem to cause any problems (yet), so I'm going to keep chugging along.

###II. Clean Data

BOOOOOO I hate this part. But I have to, and luckily I already made a function for this in EnviroStats a few months ago. From here, I'm going to get only the columns I need, and rename them. I'm going to include the addition of ProtocolType to at least my cover databases because of the 1/10 acres being stored by their event type, so I'm going to have to pull those out.

```{r}
SqueakyClean <- function(df) {
  CleanData <- df %>%
    select(Date, Ancestry, SiteID, SpeciesSymbol, SpeciesName, CommonName, SampleNumber, nValue, DDLat, DDLong) %>% 
    rename(Rank = nValue,
           Lat = DDLat,
           Lon = DDLong)
  return(CleanData)}

#This function should theoretically clean all the data. Probably going to mess up on the Tenth Acre

CleanHerbOnlyDWR <- SqueakyClean(HerbOnlyDWR)
CleanRegDWR <- SqueakyClean(RegDWR)

#As expected, there is a slight hiccup with the cover data so I'm modifying the function for it. Should just have to add the ProtocolType selection and also change nValue from Rank to Cover

AbsolutelySpotless <- function(df) {
  CleanData <- df %>%
    select(ProtocolType, Date, Ancestry, SiteID, SpeciesSymbol, SpeciesName, CommonName, SampleNumber, nValue, DDLat, DDLong) %>% 
    rename(Cover = nValue,
           Lat = DDLat,
           Lon = DDLong)
  return(CleanData)}

CleanCover <- AbsolutelySpotless(Cover)
```

###III. Getting Comp

Alrighty, now I have to get some Composition, lets see if I can replicate the process from my last project. This time, I'm going to have to do it twice per regular DWR and herbaceous DWR.

First, I have to count how much each rank was counted.

```{r}

RealRankHerbOnlyDWR <- CleanHerbOnlyDWR %>% 
  group_by(Ancestry, SiteID, Lat, Lon, Date, SpeciesSymbol, Rank) %>% 
  summarize(RankCount = n())

RealRankRegDWR <- CleanRegDWR %>% 
  group_by(Ancestry, SiteID, Lat, Lon, Date, SpeciesSymbol, Rank) %>% 
  summarize(RankCount = n())

```

Sweet, that worked great, now I'm going to move the Rank values from their rank to the value those ranks represent (1 = 7, 2 = 2, 3 = 1).

```{r}
RealRankHerbOnlyDWR <- RealRankHerbOnlyDWR %>% 
  mutate(Rank = case_when(
    Rank == 1 ~ 7,
    Rank == 3 ~ 1, #Decided not to bother typing in Rank == 2 ~ 2
    TRUE ~ Rank
  ))

RealRankRegDWR <- RealRankRegDWR %>% 
  mutate(Rank = case_when(
    Rank == 1 ~ 7,
    Rank == 3 ~ 1,
    TRUE ~ Rank
  ))
```

Important part of the whole process, I have to figure out how many total frames were read for each site. This is extremely important and will justify an extra dataframe, because since herbaceous only DWR only has perennial grasses and forbs to the exclusion of woody species and most annual plants, even the same event will have different numbers of regular DWR frames and herb only frames.

```{r}
CountedHerbOnlyDWR <- RealRankHerbOnlyDWR %>% 
  group_by(Ancestry, SiteID) %>% 
  summarize(SiteTotalFrames = sum(RankCount)/3) %>% 
  ungroup()

CountedRegDWR <- RealRankRegDWR %>% 
  group_by(Ancestry, SiteID) %>% 
  summarize(SiteTotalFrames = sum(RankCount)/3) %>% 
  ungroup()
```

Finally, we get to start making the composition, by multiplying Rank * RankCount and then adding the SiteTotalFrames and eventually Composition

```{r}
#First step, get the Rank*RankCount

HerbOnlyComp <- RealRankHerbOnlyDWR %>% 
  group_by(Ancestry, SiteID, Lon, Lat, Date, SpeciesSymbol) %>% 
  summarize(PreComp = Rank*RankCount) #PreComp is just the name I have for the value of Rank*RankCount

RegComp <- RealRankRegDWR %>% 
  group_by(Ancestry, SiteID, Lon, Lat, Date, SpeciesSymbol) %>% 
  summarize(PreComp = Rank*RankCount)

#Now I'm going to add the SiteTotalFrames to the dataframe

HerbOnlyComp <- full_join(HerbOnlyComp, CountedHerbOnlyDWR, join_by(Ancestry, SiteID))

RegComp <- full_join(RegComp, CountedRegDWR, join_by(Ancestry, SiteID))

#Finally, we get actual Composition

HerbOnlyComp <- HerbOnlyComp %>% 
  group_by(Ancestry, SiteID, Lon, Lat, Date, SpeciesSymbol, SiteTotalFrames) %>% 
  summarize(Composition = sum(PreComp))

HerbOnlyComp$Composition <- 10*HerbOnlyComp$Composition/HerbOnlyComp$SiteTotalFrames

RegComp <- RegComp %>% 
  group_by(Ancestry, SiteID, Lon, Lat, Date, SpeciesSymbol, SiteTotalFrames) %>% 
  summarize(Composition = sum(PreComp))

RegComp$Composition <- 10*RegComp$Composition/RegComp$SiteTotalFrames

```

Finally, I have to save these as CSVs so that I don't lose them.

```{r}
write_csv(HerbOnlyComp, "CleanDataTest/HerbOnlyComp.csv")
write_csv(RegComp, "CleanDataTest/RegComp.csv")
```

WHOOF I did it. Made a single dataframe with composition. Only need to make A LOT MORE FOR ALL THREE FORESTS.

Found out today that VGS is running out of funding in Sept :( not good.

###IV. Getting Cover

After doing some thinking, I have decided that I think I only need the data from Tenth Acre, and not the ones from PNF Gap, because PNF Gap will probably not have any composition to go with it. Tenth Acre is also the method we used for each method anyways, so we will continue with that.

So, what do I need? Really, only juniper cover. How do I do that? I imagine first I will have to select only data that is read with Tenth Acre. Then I will have to select only juniper cover, and then probably join that with the composition data frames so that I can compare them. I think.

Alright Step 1: select only Tenth Acre data.

```{r}
CleanCover <- filter(CleanCover, ProtocolType == "USFS R3 Tenth Acre")
```

OOF I just realized that we have recorded multiple different types of juniper codes, it could be JUNIP, JUOS, JUDE2, or JUMO. Whelp. I think I will have to figure out how to find them, and then combine them per site under JUNIP. 

Okee dokee I guess I'm doing that.

```{r}
#Filter out to only Juniperus spp.

CleanCover <- filter(CleanCover, SpeciesSymbol == "JUNIP" | SpeciesSymbol == "JUOS" | SpeciesSymbol == "JUMO" | SpeciesSymbol == "JUDE2")

#Combine them all under JUNIP code.

CleanerCover <- CleanCover %>% 
  group_by(ProtocolType, Ancestry, SiteID, Lat, Lon, Date) %>% 
  summarize(JunipCover = sum(Cover)) %>% 
  ungroup()

```

It worked, that's great. Lets see if I can add it to the Comps

```{r}
HerbOnlyComp <- full_join(HerbOnlyComp, CleanerCover, join_by(Ancestry, SiteID))

RegComp <- full_join(RegComp, CleanerCover, join_by(Ancestry, SiteID, Lon, Lat, Date))

RegComp
```

Yippee, i did it. I think now is the time for me to call it and then start analyzing the data.

OR start getting a lot more data.

Ew.